---
title: "Checking de Souza et al 2022 soybean yield data"
author: "Jim Holland"
date: '2022-09-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(pbkrtest)
library(viridis)
```

# 2020 yield data
Get yield data for 2020 soybean yield trial of transgenic vs WT contol
de Souza et al.Science 2022 Aug 19;377(6608):851-854.
doi: 10.1126/science.adc9831

```{r}
dat20 = read.csv("Q:/My Drive/Long soybean data/de Souza 2020 soybean yield data.csv") %>%
  mutate(Entry = as.factor(Entry),
         Block = as.factor(Block))
```

Plot the yield values
```{r}
ggplot(dat20, aes(x = Entry, y = Yield)) +
  geom_boxplot(aes(fill = Entry)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
linear mixed model for yield, assuming this is an RCBD
```{r}
rcbd.mod = lmer(Yield ~ (1|Block) + Entry, data = dat20)
summary(rcbd.mod, correlation = F)
```
reduced rcbd model, with no entry effects
```{r}
rcbd.mod.red = lmer(Yield ~ (1|Block), data = dat20)
```

F-test and p-value for Entry effects using Kenward Roger method

```{r}
pbkrtest::KRmodcomp(largeModel = rcbd.mod, smallModel = rcbd.mod.red)
```
Fit a fixed lm to the data, this should provide exact same results as Dataset D27
```{r}
rcbd.lm = lm(Yield ~ Block + Entry, data = dat20)
anova(rcbd.lm)
```
Yes, these results agree with their D27.  
  
Tukey HSD on genotype mean comparisons
```{r}
tuk = TukeyHSD(aov(rcbd.lm), which = "Entry")
summary(tuk$Entry[,"p adj"])
```
Tukey is too conservative here, since we don't really care about all possible comparisons.  
Really just want to test if each transgenic differs from control. That is 8 tests, so we can correct for multiple testing by using Bonferroni with 8 tests
```{r}
effects.mat = summary(rcbd.lm)$coefficients

#keep only the entry effects
effects.mat = effects.mat[grepl("Entry", rownames(effects.mat)),]
#fix up the effect names
rownames(effects.mat) = sub("Entry", "", rownames(effects.mat))
effects.mat
```
WT is the control, make all the comparisons against that.  
One of the entries has estimate value of 0.
```{r}
effects  = effects.mat[,"Estimate"]

wt = effects["WT"]

#remove the intercept and control
others = effects[!names(effects) == "WT"] 

#add in the entry with zero coefficient
zero = c(0)
names(zero) = unique(dat20$Entry)[!unique(dat20$Entry) %in% names(effects)]
others = c(others, zero)

#compute the differences between others and wt
difs = others - wt
difs
```

Compute the p-values for these comparisons. Std errors of coefficients are all the same, makes this easy
```{r}
# compute standard error of mean difference:
sed = sigma(rcbd.lm)*sqrt(2/15)
tvalues = difs/sed
pvalues = dt(tvalues, df = df.residual(rcbd.mod))
pvalues
```

The alpha 0.05 level should be divided by 8 to get Bonferroni adjusted threshold p-value for significance
```{r}
0.05/8
```
Do any of the 8 comparisons pass the Bonferroni adjustment?
```{r}
pvalues < (0.05/8)
```
They also claim to have used Benjamini-Hochberg FDR. Let's get the BH FDR adjusted p-values
```{r}
p.adjust(pvalues, method = "BH") 
```
Are any of these less than 0.05?
```{r}
p.adjust(pvalues, method = "BH") < 0.05
```
No.

# 2021 data
```{r}
dat21 = read.csv("Q:/My Drive/Long soybean data/de Souza 2021 yield data.csv") %>%
  mutate(Entry = as.factor(Entry),
         Block = as.factor(Block))
```

Plot the yield values
```{r}
ggplot(dat21, aes(x = Entry, y = Yield)) +
  geom_boxplot(aes(fill = Entry)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
The yield rankings have changed dramatically across years.
Fit these data to fixed effects ANOVA model
```{r}
rcbd.lm21 = lm(Yield ~ Block + Entry, data = dat21)
anova(rcbd.lm21)
```
No significant differences.  
Compute the means to see if they agree with Fig S12. Data are balanced, so simple means are fine:
```{r}
dat21 %>% group_by(Entry) %>%
  summarize(Mean_yield = mean(Yield))
```
Looks like their figure.


#Response of soybean yield to planting density
Data from Parks et al. 1982 TN Agric. Exp. Stn. Bull 615. Cited in Duncan 1986 Crop Science 26:584-588.

Get the data
```{r}
parks = read.csv("Q:/My Drive/Long soybean data/Parks 1983 soybean yield density data.csv")
parks.long = parks %>%
  pivot_longer(cols = c("Yield_g_m2", "Yield_g_plant"), values_to = "Yield", names_to = "Yield_type") %>%
  mutate(Yield_type = ifelse(Yield_type == "Yield_g_m2", "Yield (g/m^2)", "Yield (g/plant)"),
         Row.width = as.factor(Row.width))
```

Plot the yields per area and yield per plant as a function of population density
```{r}
density.response.plot = ggplot(parks.long, aes(x = Population, y = Yield)) + 
  geom_point(aes(colour = Row.width), size = 3) + 
  facet_wrap(facets = ~Yield_type, scales = "free") +
  scale_color_viridis(discrete = T, name = "Row width") +
  xlab("Planting density (plants per m^2)") +
  theme_minimal() +
  theme(text=element_text(size=21))

density.response.plot
```
Save the graph
```{r}
ggsave("Q:/My Drive/Long soybean data/Parks 1983 soybean yield density graph.png", density.response.plot, width = 10, height = 8, units = "in")
```

Get the data from Schutz and Bernard on G x E Interactions in the Regional Tesing of Soybean Strains. Crop Science 1967
```{r}
gxe = read.csv("Q:/My Drive/Long soybean data/Vg and Vge estimates soybean yield Shutz Bernard 1967.csv") %>%
  mutate(Vsum = Vg+ Vge,
         Genotype = Vg/Vsum,
         GxE = Vge/Vsum) %>%
  pivot_longer(cols = c('Genotype', 'GxE'), names_to = "Component", values_to = "Proportion")

gxe.plot = ggplot(gxe, aes(x = Maturity_group, y = Proportion)) +
  geom_bar(stat = "identity", aes(fill = Component), position = "dodge") +
  scale_fill_viridis(discrete = T, option = "cividis", name = "Variance\nComponent") +
  xlab("Maturity group") +
  theme_minimal()+
  theme(text=element_text(size=21))

gxe.plot
```
Write the figure to a file
```{r}
ggsave("Q:/My Drive/Long soybean data/Vg and Vge estimates soybean yield.png",gxe.plot, width = 10, height = 8, units = "in")
```

